---
output:
  md_document:
    variant: markdown_github
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path = "README-"
)
```

`bayesboot`: Easy Bayesian Bootstrap in R
========================================

The `bayesboot` package implements a function `bayesboot` that performs the Bayesian bootstrap as introduced by Rubin (1981). The implementation can both handle statistics that works on a weighted version of the data or that works on a resampled data set. 

`bayesboot` is not yet on CRAN but can be installed using [the devtools package](https://github.com/hadley/devtools):


```{r eval=FALSE}
# If you don't have devtools run this first: 
# install.packages("devtools")
devtools::install_github("rasmusab/bayesboot")
```


A simple example
---------------------

Here is a Bayesian bootstrap analysis of the mean height of the last ten American presidents:

```{r}
# Heights of the last ten American presidents in cm (Kennedy to Obama).
heights <- c(183, 192, 182, 183, 177, 185, 188, 188, 182, 185);

library(bayesboot)
b1 <- bayesboot(heights, mean)
```

The resulting posterior distribution in `b1` can now be `plot`ted and `summary`ized:

```{r fig.height= 4, fig.width=5}
summary(b1)
plot(b1)
```

While it is possible to use a summary statistics that works on a resample of the original data, it is more efficient if it's possible to use a summary statistics that works on a *reweighting* of the original dataset. Instead of using `mean` above it would be better to use `weighted.mean` like this:

```{r}
b2 <- bayesboot(heights, weighted.mean, use.weights = TRUE)
```

The result of a call to `bayesboot` will always result in a `data.frame` with one column per dimension of the summary statistic. If the summary statistic does not return a named vector the columns will be called `V1`, `V2`, etc. The result of a `bayesboot` call can be further inspected and post processed. For example:

```{r}
# Given the model and the data, this is the probability that the mean
# heights of American presidents is above the mean heights of
# American males as given by www.cdc.gov/nchs/data/series/sr_11/sr11_252.pdf
mean(c(b2[,1] > 175.9, TRUE, FALSE))
```

A more advanced example
----------------------------

A slightly more complicated example, where we do Bayesian bootstrap analysis of LOESS regression applied to the `cars` dataset on the speed of cars and the resulting distance it takes to stop. The `loess` function returns, among other things, a vector of `fitted` *y* values, one value for each *x* value in the data. These *y* values define the smoothed LOESS line and is what you would usually plot after having fitted a LOESS. Now we want to use the Bayesian bootstrap to gauge the uncertainty in the LOESS line. As the `loess` function accepts weighted data, we'll simply create a function that takes the data with weights and returns the `fitted` *y* values. We'll then plug that function into `bayesboot`:

```{r}
boot_fn <- function(cars, weights) {
  loess(dist ~ speed, cars, weights = weights)$fitted
}

bb_loess <- bayesboot(cars, boot_fn, use.weights = TRUE)
```

To plot this takes a couple of lines more:

```{r}
# Plotting the data
plot(cars$speed, cars$dist, pch = 20, col = "tomato4", xlab = "Car speed in mph", 
     ylab = "Stopping distance in ft", main = "Speed and Stopping distances of Cars")

# Plotting a scatter of Bootstrapped LOESS lines to represent the uncertainty.
for(i in sample(nrow(bb_loess), 20)) {
  lines(cars$speed, bb_loess[i,], col = "gray")
}
# Finally plotting the posterior mean LOESS line
lines(cars$speed, colMeans(bb_loess, na.rm = TRUE), type ="l",
      col = "tomato", lwd = 4)
```


More information
-----------------------

For more information on the Bayesian bootstrap see [Rubin's (1981) original paper]((https://projecteuclid.org/euclid.aos/1176345338)) and my blog post [The Non-parametric Bootstrap as a Bayesian Model](http://sumsar.net/blog/2015/04/the-non-parametric-bootstrap-as-a-bayesian-model/). The implementation of `bayesboot` is similar to the function outlined in the blog post [Easy Bayesian Bootstrap in R](http://sumsar.net/blog/2015/07/easy-bayesian-bootstrap-in-r/), but the interface is slightly different.

References
----------------

Rubin, D. B. (1981). The Bayesian bootstrap. *The annals of statistics*, 9(1), 130--134. [link to paper](https://projecteuclid.org/euclid.aos/1176345338)

